Supporting Information: Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-05-26 Ben_Manuscripts/hdphmm/cover_letter.tex
adding draft of cover letter

	%\noindent May 3\textsuperscript{rd}, 2019\\
	\noindent \today \\

@@

	\newcommand{\ManuscriptTitle}{Capturing Subdiffusive Solute Dynamics and 
	Predicting Selectivity in Nanoscale Pores with Time Series Modeling}

	We are submitting ``\ManuscriptTitle'' by Benjamin J. Coscia and Michael 
	R.  Shirts, for consideration for publication as an Article in the Journal
	of Chemical Theory and Computation. We expect this work to be of significant
	interest to researchers who study complex transport at the molecular level,
	as it describes important new approaches for estimating long time scale 
	behavior and bulk transport behavior from shorter molecular trajectories 
	exhibiting anomalous diffusion rather than Brownian motion. Please note 
	that this article has been uploaded to arXiv with differences only in 
	formatting: \href{https://arxiv.org/abs/2004.07905}{https://arxiv.org/abs/2004.07905}. \\        
%        Please note that the content of this article
%        %BJC: see if you agree with these changes
%        has been uploaded to arXiv: \href{https://arxiv.org/abs/2004.07905}{https://arxiv.org/abs/2004.07905}.
%        For submission to JCTC, we have made minor changes to the introduction,
%        abstract and conclusion in order to make it more suited for readers of 
%        the journal. We also made efforts to shorten the manuscript by moving 
%        certain discussions and figures to the supporting information. \\

    In this work, we build on the chemical intuition gained from our previous 
    study (Coscia and Shirts \textit{J. Phys.  Chem. B}, 123, 6314--6330 (2019))
    in order to formulate two different stochastic time series modeling approaches,
    %which produce trajectory realizations that share the same dynamical properties as 
    %those exhibited by solutes in our 5~$\mu$s molecular dynamics (MD) simulations.
    parameterized from 5~$\mu$s molecular dynamics (MD) simulations, capable of 
    predicting solute flux and selectivity on a macroscopic scale. 
    First, using anomalous diffusion theory, we parameterize solute motion as 
    subordinated fractional Brownian and L\'evy motion processes, based on the 
    observation that solutes experience a series of dwell times separated by
    anti-correlated hop lengths. Secondly, we add state-dependent dynamics to 
    standard Markov state modeling approaches, where we model solute motion as 
    a series of transitions between various trapped states, each with different
    dynamical properties. We demonstrate how one can project these models 
    onto macroscopic timescales in order to estimate solute flux and selectivity.
    This is among the first studies to apply time series modeling of molecular 
    simulations in order to infer behavior at long time scales and estimate bulk
    materials properties with complex anomalous diffusion mechanisms.

%	In this article, we analyze solute trajectories generated from 5~$\mu$s molecular 
%	dynamics (MD) simulations of an inverted hexagonal (H\textsubscript{II}) phase self-assembled lyotropic
%        liquid crystal (LLC) polymer membrane to parameterize two
%        different stochastic time series approaches capable of
%        predicting solute flux and selectivity on a macroscopic
%        scale. Both approaches quantitatively reproduce the mean
%        squared displacements (MSDs) exhibited by solutes in MD
%        simulations. This is among the first studies to apply time
%        series modeling of molecular simulations in order to infer
%        behavior at long time scales and estimate bulk
%        materials properties with complex anomalous diffusion
%        mechanisms.\\
%
%	This article builds on recent work of ours published in
%        J. Phys. Chem. B where we characterized transport mechanisms
%        exhibited by a series of small polar solutes in an LLC
%        Membrane (Coscia and Shirts \textit{J. Phys.  Chem. B}, 123,
%        6314--6330 (2019)).  We explore two main modeling
%        approaches. First, using anomalous diffusion theory, we
%        parameterize solute motion as subordinated fractional Brownian
%        and L\'evy motion processes, based on the observation that
%        solutes experience a series of dwell times separated by
%        anti-correlated hop lengths. Secondly, we add state-dependent
%        dynamics to standard Markov state modeling approaches, where
%        we model solute motion as a series of transitions between
%        various trapped states, each with different dynamical
%        properties.  Finally, we demonstrate how one can project models 
%        onto macroscopic timescales in order to estimate
%        solute flux and selectivity.\\

	%BJC: included phone numbers because it asked to do so in author's guide.
	\noindent Some suggestions for reviewers are:
	\begin{enumerate}

		\item Francisco Hung has used molecular simulations to study nematic liquid crystals
		and transport under confinement within various nanostructured materials 
		(Northeastern University, 617-373-8619,\\ \href{mailto:f.hung@northeastern.edu}{f.hung@northeastern.edu}).

@@

%		\item Ralf Metzler is an expert in anomalous stochastic processes, with interest in the physical
%		origins of anomalous diffusion as well as the inference of the parameters which govern these
%		processes. (Universität Potsdam, +49 331 977 5715, \href{mailto:rmetzler@uni-potsdam.de}{rmetzler@uni-potsdam.de}).

@@

	\noindent Michael R. Shirts \\
	\noindent Benjamin J. Coscia \\

@@

% LocalWords:  BJC Subdiffusive Solute Nanoscale Coscia arXiv solute lyotropic
% LocalWords:  LLC MSDs solutes Phys Chem subdiffusive evy timescales Klafter
% LocalWords:  univeristy JPCB woould nanostructured Mahesh Mahanthappa klafter
% LocalWords:  Tel Paddison

>>>>

	%\noindent May 3\textsuperscript{rd}, 2019\\
	\noindent \today

@@

	\newcommand{\ManuscriptTitle}{Statistical Inference of Transport Mechanisms and
	Long Time Scale Behavior from Time Series of Solute Trajectories in 
	Nanostructured Membranes}

	We are submitting ``\ManuscriptTitle'' by Benjamin J. Coscia, Christopher P.
	Calderon and Michael R. Shirts, for	consideration for publication as a Research
	Article in the ``Machine Learning in Physical Chemistry" special issue of the Journal 
	of Physical Chemistry, specifically in section X of % BJC: the link with section titles is an unfinished webpage
	part B of the journal. We expect this work to be of significant interest to 
	researchers who study complex dynamics in both structured and amorphous soft 
	materials. %BJC: this could be made more general. Maybe better focused for JPC though

	In this article, we apply the sticky hierarchical Dirichlet process autoregressive
	hidden Markov model (HDP-AR-HMM), a non-parametric Bayesian time series
	classification technique, to solute center-of-mass trajectories generated from
	5 $\mu s$ molecular dynamics (MD) simulations in a cross-linked inverted hexagonal
	phase lyotropic liquid crystal (LLC) membrane in order to automatically detect a
	variety of solute dynamical modes. We can use this information in order to 
	better understand the complex transport mechanisms exhibited by the solutes in 
	our simulation which will greatly increase the efficiency with which we 
	design LLC membranes for solute-specific separations.

	Our work represents a considerable advancement over our most recent study 
	characterizing the same set of solute center-of-mass time series trajectories
	(Coscia and Shirts, arXiv:2004.07905 [cond-mat.soft]). % BJC: is this the right way to cite arXiv?
	In that work, we used our qualitative understanding of solute transport 
	(Coscia and Shirts \textit{J. Phys. Chem. B}, 123, 6314--6330 (2019)) in order
	to formulate stochastic models which could be used to project solute behavior
	on macroscopic time scales, thus allowing prediction of experimentally relevant
	properties like membrane selectivity. Using the HDP-AR-HMM, we can analyze solute
	trajectories with no prior knowledge of their dynamic behavior and easily produce
	models which quantitatively describe their long time scale dynamics. Since the 
	approach is non-parametric, the complexity of our data drives the complexity of
	the model. We cluster the parameter sets in order to reduce the total space of 
	dynamical states to a small set of distinct parameters which allow us to clearly
	understand physical	mechanisms in terms of solute-membrane interactions. To our 
	knowledge, this is the first application of the HDP-AR-HMM to MD data and it is
	the first time this framework has been used to gain both mechanistic and kinetic
	insight.

	%BJC: included phone numbers because it asked to do so in author's guide.
	%BJC: need at least 5 reviewers
	\noindent Some suggestions for reviewers are:
	\begin{enumerate}

	    %BJC: these are reviewers suggested for the first stochastic transport paper
	    %BJC: not sure if bad to use twice
		\item Francisco Hung has used molecular simulations to study nematic liquid crystals
		and transport under confinement within various nanostructured materials 
		(Northeastern University, 617-373-8619,\\ \href{mailto:f.hung@northeastern.edu}{f.hung@northeastern.edu}).

@@

		%BJC: from structure paper
		\item Eric Jankowski has experience modeling soft materials using molecular dynamics
		including work done simulating structure factors from simulations, much like
		the techniques we use in our work (Boise State University, 208-426-5681,
		\href{mailto:ericjankowski@boisestate.edu}{ericjankowski@boisestate.edu})

		\item Emily Fox?  % she made the HDP-AR-HMM

@@

	\noindent Michael R. Shirts \\
	\noindent Benjamin J. Coscia \\
	\noindent Christopher P. Calderon \\

@@

% LocalWords:  BJC
